---
title: "main"
author: "Will Truskolaski"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Policy brief on the NSFG 2yr survey from 2022-2023

Split into 3 groups, female, pregnant female, and male  

# Female Data analysis
### Loading packages and data frame

Making the data a tibble cause they're the best

```{r}
library(tidyverse)
library(logger)
library(janitor)

data_female <- as_tibble(read.csv("NSFG-2022-2023-FemRespPUFData/NSFG_2022_2023_FemRespPUFData.csv"))
```

## Codebook sections
A. Demographic Characteristics; Household Roster Summary Variables (not full roster); Childhood Background
B. Pregnancy & Birth History; Adoption & Nonbiological Children
C. Marital & Relationship History; 1st Sexual Intercourse
D. Sterilizing Operations and Impaired Fecundity
E. Contraceptive History and Pregnancy Wantedness
F. Family Planning and Medical Services
G. Desires and Intentions for Future Births
H. Medical Help to Have a Baby; General & Reproductive Health
I. Health insurance; Residence and place of birth; Religion; Past and Current work (R and current spouse/partner)
J. Computer-Assisted Self-Interview (CASI) – Health status; body mass index; opposite-sex and same-sex behavior; sexual orientation and attraction; STD/HIV
risk behaviors; adverse and positive childhood events; household income and sources of income; COVID-19 experience

These are the most important
Section E: contraception
Section R: general reproduction 

We've got 1912 variables here, so we're definitely gonna have to select what we want before doing any sort of data analysis. 

```{r}
sum(is.na(data_female))
```

A lot of NA's to deal with too. 


Some important things from the user guide: 
"In addition to using sampling weights, researchers must use the design variables for the sampling stratum (VEST) and cluster (VECL) to obtain correct standard errors for their estimates." 
"Stratum Variable [=] VEST; Cluster Variable [=] VECL; 
Final post-stratified, fully adjusted case weight [=] WGT2022_2023"

