---
title: "main"
author: "Will Truskolaski"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

My personal goals for this project are to learn and showcase data pipeline skills such as cleaning and wrangling, while reinforcing statistics and visualization skills. 

1. create a brief that collects my findings and supports policy changes or initiatives. 
  a. cleaning
  b. wrangling 
  c. designing hypotheses I want to test
  d. choosing the right variables 
  e. conducting proper statistical tests with weights
  f. reviewing the results
  g. connecting this to policy initiatives
  

# Policy brief on the NSFG 2yr survey from 2022-2023

Split into 3 groups, female, pregnant female, and male  

# Female Data analysis
### Loading packages and data frame

Making the data a tibble cause they're the best

```{r}
library(tidyverse)
library(logger)
library(janitor)
library(datawizard)
library(survey)

data_female <- as_tibble(read.csv("NSFG-2022-2023-FemRespPUFData/NSFG_2022_2023_FemRespPUFData.csv"))
log_info("Data loaded")
```

## Codebook sections
A. Demographic Characteristics; Household Roster Summary Variables (not full roster); Childhood Background
B. Pregnancy & Birth History; Adoption & Nonbiological Children
C. Marital & Relationship History; 1st Sexual Intercourse
D. Sterilizing Operations and Impaired Fecundity
E. Contraceptive History and Pregnancy Wantedness
F. Family Planning and Medical Services
G. Desires and Intentions for Future Births
H. Medical Help to Have a Baby; General & Reproductive Health
I. Health insurance; Residence and place of birth; Religion; Past and Current work (R and current spouse/partner)
J. Computer-Assisted Self-Interview (CASI) – Health status; body mass index; opposite-sex and same-sex behavior; sexual orientation and attraction; STD/HIV
risk behaviors; adverse and positive childhood events; household income and sources of income; COVID-19 experience

These are the most important
Section E: contraception
Section R: general reproduction 

We've got 1912 variables here, so we're definitely gonna have to select what we want before doing any sort of data analysis. 

```{r}
sum(is.na(data_female))
```

A lot of NA's to deal with too. 


Some important things from the user guide: 
"In addition to using sampling weights, researchers must use the design variables for the sampling stratum (VEST) and cluster (VECL) to obtain correct standard errors for their estimates." 
"Stratum Variable [=] VEST; Cluster Variable [=] VECL; 
Final post-stratified, fully adjusted case weight [=] WGT2022_2023"

Finding those variables
```{r}
# Checking if there's the variables 
extract_column_names(data_female, starts_with("VEST"))
extract_column_names(data_female, starts_with("VECL"))
# Finding out which index value these columns are 
which(colnames(data_female) == "VEST")
which(colnames(data_female) == "VECL")


```


Creating the survey object 
```{r}
femsvy <- svydesign(
  id = ~VECL,
  strata = ~VEST,
  weights = ~WGT2022_2023,
  data = data_female,
  nest = TRUE
)
log_info("Survey object created")

```

Research Questions: 

Sex Education

Research Question 1: 
What is the relationship between grade when sex ed covers STDs and total number of sexual partners? 

Variables SEDSTDG w/ LIFPRTNR

```{r}
# recoding the SEDSTDG variable to categorical
femsvy$variables$SEDSTDG <- factor(femsvy$variables$SEDSTDG,
  levels = c(3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 96, 99),
  labels = c("3rd grade", "4th grade", "5th grade", "6th grade", 
             "7th grade", "8th grade", "9th grade", "10th grade",
             "11th grade", "12th grade", "1st yr college", "2nd yr college",
             "Not in school", "Don't know")
)
# DROP UNUSED FACTOR LEVELS
femsvy$variables$SEDSTDG <- droplevels(femsvy$variables$SEDSTDG)
log_info("variable recoded")
```

```{r}
STD_model <- svyglm(
  LIFPRTNR ~ SEDSTDG,
  design = femsvy,
  family = gaussian()
)
summary(STD_model)
```


```{r}
# Create data with all grade levels AS A FACTOR
pred_data <- expand_grid(
  SEDSTDG = levels(femsvy$variables$SEDSTDG)
) %>%
  mutate(
    SEDSTDG = factor(SEDSTDG, levels = levels(femsvy$variables$SEDSTDG))
  )
```

```{r}
# Get predictions
pred_fit <- predict(STD_model, newdata = pred_data)
pred_se <- SE(predict(STD_model, newdata = pred_data))
```

```{r}
# Combine into data frame
pred_df <- pred_data %>%
  mutate(
    fit = as.numeric(pred_fit),
    se = as.numeric(pred_se),
    ci_lower = fit - 1.96 * se,
    ci_upper = fit + 1.96 * se
  )
```

```{r}
# grouping variables, taking out the NA column, and making a new tibble
raw_data <- femsvy$variables |>
  mutate(SEDSTDG = fct_collapse(SEDSTDG,
    "3rd-4th grade" = c("3rd grade", "4th grade"),
    "1st-2nd yr college" = c("1st yr college", "2nd yr college")
  )) |>
  filter(!is.na(SEDSTDG)) |>
  mutate(LIFPRTNR_log = log1p(LIFPRTNR))

# Calculate means for each group
means <- raw_data |>
  group_by(SEDSTDG) |>
  summarise(mean_log = mean(LIFPRTNR_log, na.rm = TRUE), .groups = 'drop') |>
  mutate(x = as.numeric(SEDSTDG) - 0.35, xend = as.numeric(SEDSTDG) + 0.4)

# Plot the means 
ggplot(raw_data, aes(x = SEDSTDG, y = LIFPRTNR_log)) +
  geom_boxplot() +
  geom_segment(data = means, aes(x = x, xend = xend, y = mean_log, yend = mean_log),
               linetype = "dotted", color = "darkblue", linewidth = 1) +
  labs(y = "Log(Lifetime Sexual Partners)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

std_filename <- "Visualizations/RQ1_std_lifetime_partners_boxplot.pdf"
ggsave(std_filename, my_plot, width = 8, height = 6, dpi = 300)
log_info(paste0("Plot saved as ", std_filename))
```
I made a few choices for this data. The first is the grouping of 3rd with 4th grade, and 1st with 2nd year of college. This is because the groups were too small to be analyzed by themselves. This move first makes usable data out of smaller unusable groups, but it reduces the granularity of the data. The second choice I made was to perform a log transformation on the data to improve readability by reducing the impact of outliers. 

Results: 
The graphs show data that is skewed to the right. 


Research Question 2: 
How does sex ed impact STD rates?
Need to find another research question, because the variables do not exist.

```{r}
STD_CON_model <- svyglm(
  ANYMTHD ~ SEDSTDG,
  design = femsvy,
  family = gaussian()
)
summary(STD_model)
```
